Quantifying the rock-paper-scissors effect

### 2022-05-12 inst/rock_paper_scissors_monster.Rmd
fixes issue #133

```{r}
equalp.test(chess)
```

@@

(a "procyclic win" is a win that would be expected under a strong
monster).  Here we also have $3\alpha+M=0$, so to estimate the
strength of the monster we would solve

@@

(a graph is given below).  We may relax the assumption of equal
strength of the real players by using a more sophisticated hyper2
construction:

@@

```{r,testequalTAKM}
samep.test(H,c("Topalov","Anand","Karpov"))
```

@@

```{r}
specificp.gt.test(H,"M",0)
```

@@

```{r}
M <- seq(from=0.001,to=0.4,len=40)
L1 <- profsupp(H,"M",M)
L1 <- L1-max(L1)

@@

Above we see support curves for RPS monster strength using TAKM and
$T=A=K$, showing very similar results.

@@

```{r,label=testHW}
maxp(HW)
equalp.test(HW)
```

@@

```{r,label=maxpHWM}
maxp(HWM)
```

@@

```{r}
specificp.gt.test(HWM,"M",0)
```

>>>>

```{r equalptestchunk,cache=TRUE}
equalp.test(chess)
```

@@

(a "procyclic win" is a win that would be expected under a strong
monster).  Here we also have $3\alpha+M=1$, so to estimate the
strength of the monster we would solve

@@

and here we have $a=22+23+10=55$ and $b=13+12+8=33$ (a graph is given
below).  It is straightforward but not trivial to translate this into
`hyper3` idiom.  Noting that $3\alpha+M=1$ we define $\beta=3\alpha$
and then the unit sum constraint is $\beta+M=1$ and our probabilities
are

$$
\frac{\alpha+M;\alpha}{2\alpha+M   } =
\frac{\beta/3+M;\beta/3}{2\beta/3+M} =
\frac{1+2M;1-M}{2+M}
$$

(where the last expression is obtained using $\beta+M=1$).  We may set
up a log-likelihood function of $a,b$ using `hyper3` idiom as follows:

```{r,usehyper3}
TqAqK <- function(a,b){
  H3 <- hyper3(pnames=c("beta","M"))
  H3[c(beta=1/3,M=1)] %<>% inc(a  )
  H3[c(beta=1/3    )] %<>% inc(  b)
  H3[c(beta=2/3,M=1)] %<>% dec(a+b)
  return(H3)
}
H3 <- TqAqK(55,33)
loglik(c(M=0.1,beta=0.9),H3)
loglik(c(beta=0.9,M=0.1),H3)  #  NB: identical!
```

[above, we note that $\beta+M=3\alpha+M=1$, so the monster is the
fillup value with respect to $\beta$].  This is easy to maximize:

```{r,maximizeTqAqK}
optimize(function(M){loglik(c(M=M,beta=1-M),H3)},c(0.01,0.99),maximum=TRUE)
```

This gives maximum likelihood estimates of $M=0.18$ as would be
expected from the simple binomia model; the probability of a procyclic
win is about $0.625$.

We may relax the assumption of equal strength of the real
players by using a more sophisticated hyper2 construction:

@@

```{r,testequalTAKM,cache=TRUE}
samep.test(H,c("Topalov","Anand","Karpov"))
```

@@

```{r,specmtest,cache=TRUE}
specificp.gt.test(H,"M",0)
```

@@

```{r,label=plotproflik,cache=TRUE}
M <- seq(from=0.001,to=0.4,len=40)
L1 <- profsupp(H,"M",M)
L1 <- L1-max(L1)

@@

Above we see support curves for RPS monster strength using TAKM and
$T=A=K$, showing very similar results.  Note that the T=A=K curve
could have been obtained more nicely with function `TqAqK` but I
haven't got round to changing it yet.

@@

```{r,label=testHW,cache=TRUE}
maxp(HW)
equalp.test(HW)
```

@@

```{r,label=maxpHWM,cache=TRUE}
maxp(HWM)
```

@@

```{r,label=specicifgtest,cache=TRUE}
specificp.gt.test(HWM,"M",0)
```
